### 2017-02-16 proposal.tex
Added proposal.tex

\documentclass{article}

\usepackage[margin=1in]{geometry}

\begin{document}

It's a journey, not a destination.

\end{document}