### 2017-02-17 proposal.tex
Draft proposal

\documentclass[twocolumn,11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{caption}

\title{Journey: A Study of Studying Large Memory Systems}
\author{Mark Mansi, Suhas Pai, Hasnain Ali Pirzada}
\date{}

\begin{document}

\maketitle

\section{Motivation}

Recent advances in memory technologies have seen the emergence of systems with
terabytes of memory. Many current systems already have dozens or hundreds of
gigabytes of memory. Moreover, non-volatile memory technologies, which have
denser physical structures have made large memories more accessable. Thus, we
conclude that systems with terabytes of memory will become more prevelent in the
future.

Among recent advances, HP has been working on a system which they call "The
Machine", in which hundreds of cores have access to a multi-petabyte shared
memory pool (https://www.labs.hpe.com/the-machine/the-machine-distribution). Meanwhile, Intel and Micro announced their 3D Xpoint
non-volatile memory technology, which brings dense, fast, persistent memory into
the consumer and industrial realms (http://www.intel.com/content/www/us/en/architecture-and-technology/3d-xpoint-unveiled-video.html).

Much research is currently being done to study how non-volatile and large
memory systems would change system design. However, these studies encounter the
simple difficulty that most researchers do not have the funds to access these
sorts of systems for experimentation.

We note that there was a similar problem when large storage systems first became
available \cite{david, exalt}. Typically, there are two alternatives in
such a situation: trick the system into believing such hardware is available or
use a simulation (e.g. via a virtual machine or specialized tool).

Currently, though, there are few good tools avaiable to systems researchers for
understanding how systems interact with large memories. Running virtual machines
with large memories is extremely slow and limits the scale of these experiments.
Tricking the system into believing there is a large memory is more difficult
than tricking the same system into believing there is a large disk because user
programs have direct access to memory and tend to use more ad hoc in-memory
structures.

The ideal end-goal of this research is to find a general lightweight technique
or tool for studying and designing large-memory systems.

\section{Proposal}

This research will proceed in two phases. First, we will analyze and understand
the existing memory management mechanisms in Linux. In particular, we will look
at the data structures in use and the overhead of the various memory management
mechanisms. This understanding will help us to understand where current memory
management techniques are unscalable, which in turn will help us understand how
to build either an emulation tool or simulation tool for studying large memory
systems using more affordable common-sized-memory systems, such as a typical
research would have access to.

In exploring the design space, we will explore the limits of virtualization
techniques with regards to memory AND the ways in which the Linux kernel keeps
discovers and manages memory. Exploring the scalability of Linux memory
management allows us to see how well we can cause the Linux kernel to trick
itself into believing there is a large amount of memory. This is important
because we may find that this is possible but not efficient.

The second phase of the research will be to develop this general tool. We will
likely start by implementing a less-general tool which targets low-hanging
fruit. Then, we will make it more general as time permits. This second phase is
heavily influenced by what we find in the first phase, so it present our notions
of what we will do are rather shadowy.

\section{Evaluation of Solution}

In the course of our analysis phase, we will identify a number of workloads
targetted at large-memory systems. We can use these workloads again to evaluate
our solution.

\bibliography{references}{}
\bibliographystyle{plain}

\end{document}