### 2017-03-14 related_work.tex
Added some code

\documentclass[twocolumn,11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{caption}

\title{Journey: A Study of Studying Large Memory Systems}
\author{Mark Mansi, Suhas Pai, Hasnain Ali Pirzada}
\date{}

\begin{document}

\maketitle

\section{Related Work}

The emergence of new technology has in the past often required operating systems
to change. Traditionally, OS researchers and developers benchmark and test their
solutions in new environments to demonstrate their efficacy before using them in
real systems. However, researchers do not always have access to new technology
for practical reasons; for example, large memory systems are still too expensive
for most research budgets. This is not the first time that researchers have been
forced to study systems in environments they do not have access to. This section
examines past approaches to studying these systems and motivates the need for
our work.

The emergence of the internet required businesses to buy expensive machines to
keep up with demand for their services, but systems researchers did not have
access to these machines. Alameldeen et al. describe how they simulate a
multi-million dollar server using a \$2000 workstation. To do this, they go
through several rounds of scaling down, optimizing, changing, and tuning both
the benchmarks and systems they test. For example, they scale down a workload to
fit in the 1GB memory of the machine they use and increase the number of threads
to improve parallelism \cite{2kmachine}. This methodology may be accurate, but
it requires modifying the benchmarks, which is error prone and time consuming;
it can be extremely difficult to validate changes to a benchmark. In our work,
we seek to avoid changes to the experimental workloads including changes to code
or scale.

Likewise, the Quartz emulator studies the behaviour of applications on
Non-volatile Memory (NVM) while actually running on top of DRAM. It uses
existing hardware capabilities to emulate the higher latency and lower bandwidth
of most NVM technologies. While Quartz does not address the fact that the actual
memory available to the applications on NVM might be orders of magnitude higher
than on DRAM, it demonstrates that new technologies may be emulated using
existing technology \cite{quartz}.

The Simics simulator uses demand paging for simulated environments, which means
that memory is only allocated when it is used. As long as the working set of the
target system fits in the host memory, performance will be tolerable.  However,
in our work, we wish to study OS behavior under workloads that actually use all
available memory. Thus, the usefulness of demand paging is greatly reduced. For
these workloads, the performance of simulators like Simics tends to be
impractical \cite{simics}. We do not explore simulation further because it can
be orders of magnitude slower than native execution, making it infeasible for
studying large workloads or systems \cite{2kmachine}.

A similar problem was encountered by researchers studying both large storage
drives and large distributed storage systems. David is a system that allows
storage and big data researchers to run large benchmarks requiring terabytes of
storage using off the shelf storage devices (which at that time were too small).
David creates a compressed version of the file system by physically storing only
metadata and discarding the contents of files. Reads to the disk causes David to
generate data on the fly. This decision choice is based on the observation that
most benchmarking frameworks do not care about the actual content of the files,
and that most of the storage capacity of a drive tends to be data rather than
metadata \cite{david}. The Exalt system uses a similar methodology for
large-scale distributed storage systems \cite{exalt}. This methodology provides
a promising direction for large-memory system studies. One may consider ignoring
the contents of a process’s heap and only storing kernel data structures and a
process’s code and stack segments. However, generating heap contents on the fly
is more difficult than generating disk contents because of the common use of
custom data structures. Moreover, on Linux, the kernel data structures for 1TB
of memory may also exceed the size of physical memory on present systems
\cite{simics}.

A virtualized environment can be used to provide a guest system with more memory
than is physically available to the host. A study of the limits of the KVM
hypervisor found that there is no fundamental limit to the size of guest
physical memory other than the hardware address width. However, currently, a
Linux host running KVM will require the guest memory to be backed by host
physical memory or host swap space \cite{ibmkvm}.  This means that when the
virtual machine uses the whole amount of memory allocated to it, the host can
swap pages to disk. The resulting poor performance can cause inaccurate
performance measurements when running benchmarks.

Prefetching pages from swap space can offer a way to mitigate the overhead of
memory overcommitment. When there is significant memory pressure, even pages
which are likely to be accessed soon are swapped out to disk. They are faulted
back into memory when accessed, resulting in significant performance cost.
Charm++ uses a programming model where computations can be scheduled by the
language runtime. A designated thread can prefetch pages required by a
computation, averting page faults. Although this approach is effective in
avoiding overhead, it requires applications to be written with a particular
programming language and so is not applicable to most benchmarks without
changing them \cite{charmpp}.

Gupta et al. built an emulator for high speed networks using a technique called
time dilation, which slows down the OS clock to make it appear that external
events are occurring faster. This allows the system to emulate network links
with speeds that are currently not available. The implementation is based on the
VMs and Xen hypervisor; Xen delivers the timer interrupts to the guest at a
lower rate than hardware hence  slowing down the guest’s clock \cite{timedil}.
We may use a similar approach to slow down system time  while paging in and out
large portions of memory. This will allow us to make the system believe that it
is reading data from the memory while actually most of that data is being read
from the disk.

Finally, there have been studies which look at the performance overheads
associated with current implementations of virtual memory and suggest mechanisms
to mitigate them. RadixVM tries to overcome performance issues in highly
concurrent workloads due to serialization of memory management operations on
kernel data structures \cite{radixvm}. This work demonstrates that many parts of
existing memory management schemes are not scalable to larger systems.
Similarly, in our work, we wish to examine scalability limitations of memory
management in the Linux kernel. Other studies have suggested that \texttt{struct
page}, \texttt{struct vm\_area\_struct}, and page tables tend to comprise a
large portion of memory management overhead \cite{simics}.

\bibliography{references}{}
\bibliographystyle{plain}

\end{document}